---
title: "Practical Machine Learning Course Project"
output:
  html_document: default
  pdf_document: default
---


### Wojciech Starosta

#### April 5, 2018

##Executive Summary

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

##Libraries and data
The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r lib, message=FALSE, warning=FALSE}
library(caret); library(rpart);library(ggplot2);library(randomForest); library(rpart.plot)
```

```{r data, message=FALSE, warning=FALSE}
#downloading data
url.train <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
url.test <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training <- read.csv(url(url.train), na.strings = c("NA", "", "#DIV0!"))
testing <- read.csv(url(url.test), na.strings = c("NA", "", "#DIV0!"))
training <- training[, colSums(is.na(training)) == 0]
testing <- testing[, colSums(is.na(testing)) == 0]

# Delete variables that are not related 
training <- training[, -c(1:7)]
testing <- testing[, -c(1:7)]

# partioning the training set into two different dataset with 60/40 share
traningPartition <- createDataPartition(training$classe,  p = 0.6, list = F)
trainingSet <- training[traningPartition, ]
testingSet <- training[-traningPartition, ]
dim(training); dim(testingSet)
```

##Model 1 - Decision trees
```{r dc, message=FALSE, warning=FALSE}
decisionTreeModel <- rpart(classe ~ ., data = trainingSet, method = "class")
decisionTreePrediction <- predict(decisionTreeModel, testingSet, type = "class")

# Plot Decision Tree
#rpart.plot(decisionTreeModel, main = "Decision Tree", under = T, faclen = 0)

# Testing results
confusionMatrix(decisionTreePrediction, testingSet$classe)
```

##Model 2 - Random Forest
```{r rf, message=FALSE, warning=FALSE}
forestmodel <- randomForest(classe ~ ., data=trainingSet, method="class")
predictionforest <- predict(forestmodel, testingSet, type="class")

# Testing results
confusionMatrix(predictionforest, testingSet$classe)
```

##Conclusion
The goal of this project was to predict the manner in which participants did the exercise. Data was divided into training and testing set and two kind of models were builded to predict output on second set. First was Decision Tree which gives accuracy on 0.7355 level with kappa value equal to 0.664. Second Random Forest was applied with accuracy 0.9934 and kappa 0.9916. It can be easily seen that Random Forest is way much better than Decision Trees.

This project is reproducible and was done with the following environment:

```{r sessioninfo, message=FALSE, warning=FALSE}
sessionInfo()
```